Influence of School Closure on COVID-19 Contaminations in Ireland and Repercussions across Age Groups

### 2022-01-18 manuscript/manuscript.Rmd
sent to edgar

output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("output"),
      output_file = glue::glue("{xfun::sans_ext(input)}.pdf"),
      envir = globalenv()
    )
  })
---

@@

# options ----------------------------------------------------------------------
set.seed(123) # Seed for random number generation
options(scipen = 999) # disable sci number format
opts_chunk$set(
  cache.extra = rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  echo = FALSE, 
  fig.retina = 3
  )

@@

\begin{align}
  T_{J \rightarrow I}(k,l) = \sum_{i,j} p\left(i_{t+1}, i_t^{(k)}, j_t^{(l)}\right) \cdot log \left(\frac{p\left(i_{t+1}| i_t^{(k)}, j_t^{(l)}\right)}{p\left(i_{t+1}|i_t^{(k)}\right)}\right),
\end{align}

\noindent where $T_{J\rightarrow I}$ consequently measures the influence of the change dynamic from an age group $J$ to another age group $I$ (Eq 3).

@@

In order to evaluate the shape of the trend in the numbers of Covid-19 cases reported after the three school closure longer than 14 consecutive days, a GAM was fitted taking into account numbers for 1 to 4, 5 to 14, and 15 to 24 year old. The obtained results satisfy the requirements to fit this model which explains `r scales::percent(gam_summary$dev.expl, accuracy = 0.1)` of the deviance in Covid-19 cases. These results show no significant change in the evolution of cases for 1 to 4 year old (`r gam_results["s(date):1to4"]`) as well as for 5 to 14 year old (`r gam_results["s(date):5to14"]`). However, the smooth term of the GAM is significant for 15 to 24 year old (`r gam_results["s(date):15to24"]`), which indicates that the evolution of cases is not linear and, therefore, the school closure had a significant impact on the evolution of cases.

@@

Discussion
===================

@@

This indicate mix results, while the school closure was ineffective in child care settings and primary school, the closure of secondary schools and colleges had an impact on Covid-19 transmission.

Conclusion
===================

>>>>

output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    keep_tex: false
  bookdown::word_document2:
    default
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("output"),
      output_file = c(glue::glue("{xfun::sans_ext(input)}.pdf"), glue::glue("{xfun::sans_ext(input)}.docx")), 
      output_format = "all",
      envir = globalenv()
    )
  })
---

@@

# options ----------------------------------------------------------------------
set.seed(123) # Seed for random number generation
options(scipen = 999) # disable sci number format
opts_chunk$set(
  cache.extra = rand_seed, 
  cache = TRUE,
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  echo = FALSE, 
  fig.retina = 3
  )

@@

\begin{align}
  T_{X \rightarrow Y}(k,l) = \sum_{x,y} p\left(x_{t+1}, x_t^{(k)}, y_t^{(l)}\right) \cdot log \left(\frac{p\left(x_{t+1}| x_t^{(k)}, y_t^{(l)}\right)}{p\left(x_{t+1}|x_t^{(k)}\right)}\right),
\end{align}

\noindent where $T_{X\rightarrow Y}$ consequently measures the influence of the change dynamic from an age group $X$ to another age group $Y$ (Eq 3).

@@

In order to evaluate the shape of the trend in the numbers of Covid-19 cases reported after the three school closure longer than 14 consecutive days, a GAM was fitted taking into account numbers for 1 to 4, 5 to 14, and 15 to 24 year old. The obtained results satisfy the requirements to fit this model which explains `r scales::percent(gam_summary$dev.expl, accuracy = 0.1)` of the deviance in Covid-19 cases. These results show no significant change in the evolution of cases for 1 to 4 year old (`r gam_results["s(date):1to4"]`) as well as for 5 to 14 year old (`r gam_results["s(date):5to14"]`). However, the smooth term of the GAM is significant for 15 to 24 year old (`r gam_results["s(date):15to24"]`), which indicates that the evolution of cases is not linear and, therefore, the school closure had a significant impact on the evolution of cases (Figure \@ref(fig:gam)).

@@

The results of the Transfer Entropy calculations between age groups are reported Figure \@ref(fig:te). The first observation is the absence of symmetry between influencing age groups (i.e., $X$) and influenced age groups (i.e., $Y$). Indeed, the change in Covid-19 cases in some age groups are influenced by other age groups but they are not reciprocally influencing these age groups. This is the case of 15 to 24 years old who appears to be contaminated by all age groups until 84 but there is no significant support that they are contaminating groups older than 34 which should include their parents and relatives.

Discussion & Conclusion
===================

@@

This indicate mix results, while the school closure was ineffective in child care settings and primary school, the closure of secondary schools and colleges had an impact on Covid-19 transmission. This impact is probably responsible for the absence of reciprocity in the transmission of 15 to 24 age group. By drastically reducing their social contacts in secondary schools and colleges, this cohort has more been the target of Covid-19 contaminations rather than the vector of Covid-19.
